### 2020-05-30 writeup/experiments.tex
typos
merge
sec 3

\section{Practical Implications for Multi-Task Learning}

@@

\subsection{Detecting and Improving Negative Effects}

\textbf{A Metric to Determine Positive or Negative Transfer}
We propose a simple metric to determine whether multi-task learning performs better than single-task learning.
The metric is as follows.
First, we train 2 single-task models on 2 separate tasks.
We take the prediction accuracy of the two tasks.
Let $\tau$ be a fixed threshold.
If the accuracy of the source task is higher than the accuracy of the target task by $\tau$, then we predict positive transfer.
On the other hand, if the accuracy of the source task is lower than the accuracy of the target task by $\tau$, then we predict negative transfer.

Table \ref{tab:mtl_better_than_stl} shows the results on a sentiment analysis and an image classification task.

@@

@@

\subsection{Validating Theoretical Implications}

\textbf{Causes of negative transfer.}

\textbf{Benefits of multi-task learning.}

\textit{Improving labeled data efficiency.}
We measure the data efficiency ratio on text classification tasks.
We find that by performing multi-task learning, only 40$\%$ of the data is needed to achieve comparable performance to single-task learning over all six tasks.

\textit{Transferring from a high accuracy to low accuracy.}

\begin{table}
	\begin{center}
		\begin{tabular}{c c c c c}
		\toprule
		\multirow{2}{*}{{\bf Models}} & \multicolumn{2}{c}{\begin{minipage}{1.1in}\begin{center}
		MR, SST, SUBJ, CR, MPQA, TREC\end{center}\end{minipage}} & \multicolumn{2}{c}{\begin{minipage}{1.1in}\begin{center}MR, SST, SUBJ, CR, MPQA\end{center}\end{minipage}} \\
		\cmidrule(lr){2-3} \cmidrule(lr){4-5}
		& {\bf Stanford} & {\bf Alignment} & {\bf Stanford} & {\bf Alignment} \\
		\midrule
		{\bf MLP}  & > 100\% & 39\% & 25\% & 25\% \\
		{\bf LSTM} & 36\% & 36\% & 28\% & 25\% \\
		% {\bf CNN}  & 76\% & - & 32\% & -\\
		\bottomrule
		\end{tabular}
	\end{center}
	\caption{Taskonomy experiment.}
	\label{tab:taskonomy}
	\end{table}

>>>>

\section{Experiments}

@@

\subsection{Improving training efficiency in MTL}

\subsection{Validating Algorithmic Implications}

\textbf{A Metric to Determine Positive or Negative Transfer}
Table \ref{tab:mtl_better_than_stl} shows the results on a sentiment analysis and an image classification task.

@@

\textbf{Improving labeled data efficiency.}
We measure the data efficiency ratio on text classification tasks.
We find that by performing multi-task learning, only 40$\%$ of the data is needed to achieve comparable performance to single-task learning over all six tasks.

@@

%\subsection{Validating Theoretical Implications}

%\textbf{Contributing causes of negative transfer.}

%\textbf{Implications for positive learning.}

%\textit{Transferring from a high accuracy to low accuracy.}%%%%%%%%

\begin{table}
	\begin{center}
		\begin{tabular}{c c c c c}
		\toprule
		\multirow{2}{*}{{\bf Models}} & \multicolumn{2}{c}{\begin{minipage}{1.1in}\begin{center}
		MR, SST, SUBJ, CR, MPQA, TREC\end{center}\end{minipage}} & \multicolumn{2}{c}{\begin{minipage}{1.1in}\begin{center}MR, SST, SUBJ, CR, MPQA\end{center}\end{minipage}} \\
		\cmidrule(lr){2-3} \cmidrule(lr){4-5}
		& {\bf Standard} & {\bf Alignment} & {\bf Standard} & {\bf Alignment} \\
		\midrule
		{\bf MLP}  & > 100\% & 39\% & 25\% & 25\% \\
		{\bf LSTM} & 36\% & 36\% & 28\% & 25\% \\
		% {\bf CNN}  & 76\% & - & 32\% & -\\
		\bottomrule
		\end{tabular}
	\end{center}
	\caption{Taskonomy experiment.}
	\label{tab:taskonomy}
	\end{table}

### 2020-05-30 writeup/setup.tex
sec 3

\section{Preliminaries}\label{sec_setup}

We describe the bias-variance tradeoff of our setting more formally.
Recall that we have $t$ labeled tasks available, denoted by $(X_1, Y_1), (X_2, Y_2), \dots, (X_t, Y_t)$, where $X_i\in\real^{n_i\times p}$ and $Y_i\in\real^{n_i}$ for $1\le i\le t$.
%Following \cite{HMRT19,BLLT20}, we assume that for each task $i = 1,2,\dots,t$,  every feature vector is generated as $x = \Sigma_i^{1/2} z$, where $z\in\real^p$ is a random vector with i.i.d. entries of mean zero and unit variance and $\Sigma_i\in\real^{p\times p}$ is a positive semidefinite matrix.
Without loss of generality, let the $t$-th task denote the target task.
For an estimator $\hat{\beta}\in\real^p$, we define the out-of-sample (prediction) loss as
	\begin{align*}
		\te_t(\hat{\beta}) \define \exarg{z}{\exarg{\varepsilon_t}{({(\Sigma_t^{1/2} z)}^{\top}\hat{\beta} - {(\Sigma_t^{1/2})}^{\top}\beta_t)^2}}
		= \exarg{\varepsilon_t}{(\hat{\beta} - \beta_t)^{\top}\Sigma_t(\hat{\beta} - \beta_t)}.
	\end{align*}
The single-task estimator $\hat{\beta}_t^{\STL}$ is given by $(X_t^{\top}X_t)^{-1}X_t^{\top}Y_t$.
The bias-variance trade-off \cite{HTF09} says
	\[ \te_t(\hat{\beta}) =
		\bignorm{\exarg{\varepsilon_t}{\hat{\beta}} - \beta_t}^2 + \exarg{\varepsilon_t}{\bignorm{\hat{\beta} - \exarg{\varepsilon_t}{\hat{\beta}}}^2}. \]
In order to study the trade-off between model-shift bias and variance reduction, we need tight concentration bounds to quantify both effects.
For this purpose, we consider the high-dimensional regime where $n_i$ is a fixed constant $\rho_i > 1$ times $p$ for every $1\le i\le t$, and $p$ is large.
%Recall that $n_i = \rho_i \cdot p$ and we assume $\rho_i > 1$ is a fixed constant for every $1\le i\le t$.

@@

To illustrate our intuition, we begin by considering the setting of two tasks with general covariance matrices.
Recall that $\hat{\beta}_t^{\MTL}$ is defined as $BW_t$ after solving equation \eqref{eq_mtl}.
We decompose the test error of $\hat{\beta}_{t}^{\MTL}$ on the target task into two parts (to be derived in Appendix \ref{app_proof_sec3}) as follows
\begin{align}
	\te_t(\hat{\beta}_t^{\MTL}) =& ~ \hat{v}^2 \bignorm{\Sigma_2^{1/2} (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1 (\beta_1 - \hat{v}\beta_2)}^2 \label{eq_te_model_shift} \\
	&+ \sigma^2\cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}, \label{eq_te_var}
\end{align}
where $\hat{v} = W_2 / W_1$ denotes the ratio of the output layer weights.
%Hence the bias of $\te_t(\hat{\beta}_t^{\STL})$ is zero and its test error is equal to variance, given by $$.

\textbf{Notations.}
When there is no ambiguity, we drop the subscript $t$ from $\te_t(\hat{\beta}_t^{\MTL})$ to $\te(\hat{\beta}_t^{\MTL})$ for simplicity.
We refer to the first task as the source task when there are only two tasks.
We call $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ the covariate shift matrix.
%Same for $\hat{\beta}_t^{\STL}$.

\section{A Technical Tool to Quantify the Bias-Variance Trade-off}
\label{sec_main}

We develop a technical tool to derive $\te(\hat{\beta}_t^{\MTL})$ that only depends on the qualities of task data such as data sizes and covariance matrices, for two tasks with general covariances.
Then, we extend the result to more than two tasks that share the same features but have different labels.
Finally, we show a sharp bias-variance tradeoff for transfer learning settings.

@@

\begin{lemma}[Informal statement of Lemma \ref{lem_cov_shift}]\label{lem_cov_shift_informal}
	Suppose $X_1=Z_1\Sigma_1^{1/2}\in \R^{n_1\times p}$ and $X_2=Z_2\Sigma_2^{1/2}\in \R^{n_2\times p}$ with $\rho_1=n_1/p>1$ and $\rho_2=n_2/p>1$.
	Let $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and $\lambda_1, \lambda_2, \dots, \lambda_p$ be the singular values of $M^{\top}M$ in descending order.
	For any constant $\e>0$, w.h.p. over the randomness of $X_1, X_2$, we have that
	\begin{align*}
		\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2} = \frac{1}{\rho_1+\rho_2}\cdot \frac1p\bigtr{ (a_1 \Sigma_1 + a_2\Sigma_2)^{-1} \Sigma_2} +\bigo{\|\Sigma_2\| p^{-1/2+\epsilon}},
	\end{align*}
where $(a_1, a_2)$ is the solution to the following deterministic equations:
	\begin{align*}
		a_1 + a_2 = 1- \frac{1}{\rho_1 + \rho_2},\quad a_1 + \frac1{\rho_1 + \rho_2}\cdot \frac{1}{p}\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2} = \frac{\rho_1}{\rho_1 + \rho_2}.
	\end{align*}
\end{lemma}

\subsection{Applications to Multi-Task and Transfer Learning}

\textbf{Multi-task learning.} applies to the setting of two tasks where their covariance matrices may be arbitrarily different.
The above result also applies to the setting of more than two tasks with the same covariates.
%We extend the above result to any number of tasks that have the same covariates.
%In this section we consider the setting with $k$ many that have the same covariates.
Since the tasks all have the same number of datapoints and covariance matrix, the trade-off between model shift bias and variance will be captured by their task models $\set{\beta_i}_{i=1}^k$.
%For this setting, we derive solutions for the multi-task training and the transfer learning setting that match our insights qualitatively from Section \ref{sec_denoise}.
%Let $B^{\star} = [\beta_1, \beta_2, \dots, \beta_k] \in\real^{p\times k}$ denote the underlying task model parameters.
The formal statement is stated in Theorem \ref{thm_many_tasks} and its proof can be found in Appendix \ref{app_proof_many_tasks}.
%The technical crux of our approach is to derive the asymptotic limit of $\te(\hat{\beta}_t^{\MTL})$ in the high-dimensional setting, when $p$ approaches infinity.
We derive a precise limit of $\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}$, which is a deterministic function that only depends on $\Sigma_1, \Sigma_2$ and $n_1/p, n_2/p$ (see Lemma \ref{lem_cov_shift} in Appendix \ref{app_proof_main} for the result).
Based on the result, we show how to determine positive versus negative transfer as follows.

\begin{theorem}[Informal statement of Theorem \ref{thm_model_shift}]\label{thm_main_informal}
	Let $X_i \in\real^{n_i \times p}$ and $Y_i = X_i\beta_i + \varepsilon_i$, for $i = 1, 2$.
	Suppose that $n_1 = \rho_1 p$ and $n_2 = \rho_2 p$, where $\rho_1>1$ and $\rho_2 >1$ are fixed constants.
	There exists two deterministic functions $\Delta_{\beta}$ and $\Delta_{\vari}$ and a small deterministic error $\delta$ that only depend on $\set{\hat{v}, \Sigma_1^{}, n_1, n_2, \beta_1, \beta_2}$ such that
	\begin{itemize}
		\item If $\Delta_{\vari} - \Delta_{\beta} \ge \delta$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\Delta_{\vari} - \Delta_{\beta} \le \delta$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\end{itemize}
\end{theorem}

Theorem \ref{thm_main_informal} shows nearly tight bounds on the trade-off between model-shift bias and variance reduction.
%determined by the covariate shift matrix and the model shift.
The bounds get tighter and tighter as $\rho_1$ increases.
While the general form of $\Delta_{\vari}$ and $\Delta_{\beta}$ can be quite complex, we will show that they provide nice interpretation for simplified settings later on in Section \ref{sec_insight}.
A formal version of Theorem \ref{thm_main_informal} is presented in Theorem \ref{thm_model_shift} and its proof is presented in Appendix \ref{app_proof_main}.

>>>>

\section{A Technical Tool to Analyze the Bias-Variance Trade-off}
\label{sec_main}

We first describe the bias-variance tradeoff in our setting more formally.
Then, we develop a key lemma for analyzing the tradeoff using random matrix theory.
%We develop a technical tool to derive $\te(\hat{\beta}_t^{\MTL})$ that only depends on the qualities of task data such as data sizes and covariance matrices, for two tasks with general covariances.
%Then, we extend the result to more than two tasks that share the same features but have different labels.
Finally, we apply the key lemma to provide a sharp analysis of the bias-variance tradeoff of multi-task and transfer learning in our setting.

\subsection{The Bias-Variance Tradeoff}\label{sec_setup}

Recall that we have $t$ labeled tasks available, denoted by $(X_1, Y_1), (X_2, Y_2), \dots, (X_t, Y_t)$, where $X_i\in\real^{n_i\times p}$ and $Y_i\in\real^{n_i}$ for $1\le i\le t$.
%Following \cite{HMRT19,BLLT20}, we assume that for each task $i = 1,2,\dots,t$,  every feature vector is generated as $x = \Sigma_i^{1/2} z$, where $z\in\real^p$ is a random vector with i.i.d. entries of mean zero and unit variance and $\Sigma_i\in\real^{p\times p}$ is a positive semidefinite matrix.
Without loss of generality, let the $t$-th task denote the target task.
For an estimator $\hat{\beta}\in\real^p$, we define the out-of-sample (prediction) loss as
	\begin{align*}
		\te_t(\hat{\beta}) \define \exarg{z}{\exarg{\varepsilon_t}{({(\Sigma_t^{1/2} z)}^{\top}\hat{\beta} - {(\Sigma_t^{1/2})}^{\top}\beta_t)^2}}
		= \exarg{\varepsilon_t}{(\hat{\beta} - \beta_t)^{\top}\Sigma_t(\hat{\beta} - \beta_t)}.
	\end{align*}
The single-task estimator $\hat{\beta}_t^{\STL}$ is given by $(X_t^{\top}X_t)^{-1}X_t^{\top}Y_t$.
The bias-variance trade-off \cite{HTF09} says
	\[ \te_t(\hat{\beta}) =
		\bignorm{\exarg{\varepsilon_t}{\hat{\beta}} - \beta_t}^2 + \exarg{\varepsilon_t}{\bignorm{\hat{\beta} - \exarg{\varepsilon_t}{\hat{\beta}}}^2}. \]
In order to study the trade-off between model-shift bias and variance reduction, we need tight concentration bounds to quantify both effects.
For this purpose, we consider the high-dimensional regime where $n_i$ is a fixed constant $\rho_i > 1$ times $p$ for every $1\le i\le t$, and $p$ is large.
%Recall that $n_i = \rho_i \cdot p$ and we assume $\rho_i > 1$ is a fixed constant for every $1\le i\le t$.

@@

\textbf{Notations.}
When there is no ambiguity, we drop the subscript $t$ from $\te_t(\hat{\beta}_t^{\MTL})$ to $\te(\hat{\beta}_t^{\MTL})$ for simplicity.
We refer to the first task as the source task when there are only two tasks.
We call $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ the covariate shift matrix.

@@

To illustrate our intuition, we begin by considering the setting of two tasks with general covariance matrices.
Recall that $\hat{\beta}_t^{\MTL}$ is defined as $BW_t$ after solving equation \eqref{eq_mtl}.
We decompose the test error of $\hat{\beta}_{t}^{\MTL}$ on the target task into two parts (to be derived in Appendix \ref{app_proof_sec3}) as follows
\begin{align}
	\te(\hat{\beta}_t^{\MTL}) =& ~ \hat{v}^2 \bignorm{\Sigma_2^{1/2} (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1 (\beta_1 - \hat{v}\beta_2)}^2 \label{eq_te_model_shift} \\
	&+ \sigma^2\cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}, \label{eq_te_var}
\end{align}
where $\hat{v} = W_2 / W_1$ denotes the ratio of the output layer weights.
The scaling term $\hat{v}$ corresponds to the intuition that the shared module $B$ learns a subspace for both tasks.
The output layer of each task scales the direction of $B$ suitably to fit the task.

We term equation \eqref{eq_te_model_shift} as \textit{model-shift bias}, which captures how similar $\beta_1$ and $\beta_2$ are.
This part introduces a negative effect to multi-task learning.
Next, it is not hard to verify that equation \eqref{eq_te_var}, the variance of $\hat{\beta}_t^{\MTL}$, is always smaller than the variance of $\hat{\beta}_t^{\STL}$.
This part introduces a positive variance reduction effect to performing multi-task learning.
Hence, whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$ is determined by the tradeoff between two effects:
(i) the positive effect from variance reduction;
(ii) the negative effect from model shift bias.
In order to analyze the tradeoff, we present the following lemma, which provides a tight bound for equation \eqref{eq_te_var}.

\begin{lemma}[Informal statement of Lemma \ref{lem_cov_shift}]\label{lem_cov_shift_informal}
	In the setting of two tasks, let $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and $\lambda_1, \lambda_2, \dots, \lambda_p$ be the singular values of $M^{\top}M$ in descending order.
	For any constant $\e>0$, w.h.p. over the randomness of $X_1, X_2$, we have that
	\begin{align*}
		\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2} = \frac{1}{\rho_1+\rho_2}\cdot \frac1p\bigtr{ (a_1 \Sigma_1 + a_2\Sigma_2)^{-1} \Sigma_2} +\bigo{\|\Sigma_2\| p^{-1/2+\epsilon}},
	\end{align*}
where $(a_1, a_2)$ is the solution to the following deterministic equations:
	\begin{align*}
		a_1 + a_2 = 1- \frac{1}{\rho_1 + \rho_2},\quad a_1 + \frac1{\rho_1 + \rho_2}\cdot \frac{1}{p}\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2} = \frac{\rho_1}{\rho_1 + \rho_2}.
	\end{align*}
\end{lemma}

As a remark, when there is only the target task, $\rho_1$ equals zero.
Hence $a_2 = 1 - 1/ \rho_2, a_1 = 0$ and Lemma \ref{lem_cov_shift_informal} states that $\tr[(X_2^{\top}X_2)^{-1}\Sigma_2] = \sigma^2 / (\rho_2 - 1)$, which is to a well-known result in random matrix theory \cite{S07}.

\subsection{A Sharp Analysis of the Tradeoff for Multi-Task and Transfer Learning}

\textbf{Multi-task learning.}
We use Lemma \ref{lem_cov_shift_informal} to provide a sharp bias-variance tradeoff for two settings:
(i) two tasks with general covariance matrices;
(ii) any number of tasks that have the same features.
This setting is prevalent in applications of multi-task learning to image classification, where there are multiple prediction labels/tasks for every image \cite{chexnet17,EA20}.
We state our result for two tasks as follows.
%The formal statement is stated in Theorem \ref{thm_many_tasks} and its proof can be found in Appendix \ref{app_proof_many_tasks}.
%The technical crux of our approach is to derive the asymptotic limit of $\te(\hat{\beta}_t^{\MTL})$ in the high-dimensional setting, when $p$ approaches infinity.
%We derive a precise limit of $\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}$, which is a deterministic function that only depends on $\Sigma_1, \Sigma_2$ and $n_1/p, n_2/p$ (see Lemma \ref{lem_cov_shift} in Appendix \ref{app_proof_main} for the result).
%Based on the result, we show how to determine positive versus negative transfer as follows.

\begin{theorem}[Informal statement of Theorem \ref{thm_model_shift}]\label{thm_main_informal}
	For the setting of two tasks, let $\delta > 0$ be a desired error margin and $\rho_1 > ??$.
	There exists two deterministic functions $\Delta_{\beta}$ and $\Delta_{\vari}$ that only depend on $\set{\hat{v}, \Sigma_1^{}, n_1, n_2, \beta_1, \beta_2}$ such that
	\squishlist
		\item If $\Delta_{\vari} - \Delta_{\beta} \ge \delta$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\Delta_{\vari} - \Delta_{\beta} \le \delta$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\squishend
\end{theorem}

Theorem \ref{thm_main_informal} shows nearly tight bounds on the trade-off between model-shift bias and variance reduction.
%determined by the covariate shift matrix and the model shift.
The bounds get tighter and tighter as $\rho_1$ increases.
While the general form of $\Delta_{\vari}$ and $\Delta_{\beta}$ can be quite complex, we will show that they provide nice interpretation for simplified settings later on in Section \ref{sec_insight}.
The proof of Theorem \ref{thm_main_informal}  is presented in Appendix \ref{app_proof_main}.
The result for more than two tasks is deferred to Appendix \ref{app_proof_many_tasks}.

### 2020-05-30 writeup/transfer.tex
sec 3

\section{Theoretical Implications}
\label{sec_insight}

Based on Theorem \ref{thm_main_informal}, we provide precise interpretations of many empirical phenomena in multi-task learning.
In Section \ref{sec_similarity}, we explain three contributing causes of negative transfer, including model dissimilarity, large source data size and covariate shift.
In Section \ref{sec_benefit}, we describe two implications for positive transfer, including labeled data efficiency and de-noising.

\subsection{Contributing Causes of Negative Transfer}\label{sec_similarity}

@@

\textbf{Cause 1.}
We measure model dissimilarity as $\norm{\beta_1 - \beta_2}^2$, which is the distance between source and target in the isotropic model.
We derive a sharp threshold when positive transfer transitions to negative transfer, as model dissimilarity increases.
%Based on Theorem \ref{thm_model_shift}, we derive the transition threshold in the following proposition.
Define the following functions
\begin{align*}
	\Psi(\beta_1, \beta_2) = {\ex{\bignorm{\beta_1 - \beta_2}^2}} / {\sigma^2},  \Phi(\rho_1, \rho_2) = \frac{(\rho_1 + \rho_2 - 1)^2}{\rho_1 (\rho_1 + \rho_2) (\rho_2 - 1)}.
\end{align*}
Let $\alpha_{-} = (1 - \rho_1^{-1/2})^4$ and $\alpha_{+} = (1 + \rho_1^{-1/2})^4$.

@@

Proposition \ref{prop_dist_transition} simplifies Theorem \ref{thm_main_informal} in the isotropic model, allowing for a more explicit statement of the bias-variance tradeoff.
We remark that the constants $1/2$ and $2$ in the statement above are chosen for ease of presentation.
In general we can close their gap by increasing $\rho_1$.

We describe an overview of the proof of Theorem \ref{thm_main_informal}.
By comparing the bias and variance of $\te_t(\hat{\beta}_t^{\MTL})$ to $\te_t(\hat{\beta}_t^{\STL})$, we observe the following two effects, respectively.
%We begin by observing that the test error of $\hat{\beta}_t^{\MTL}$ consists of two parts.
We term equation \eqref{eq_te_model_shift} as \textit{model-shift bias}, which captures how similar $\beta_1$ and $\beta_2$ are.
This part introduces a negative effect to multi-task learning.
The scaling term $\hat{v}$ corresponds to the intuition that the shared module $B$ learns a subspace for both tasks.
The output layer of each task scales the direction of $B$ suitably to fit the task.
Next, it is not hard to verify that equation \eqref{eq_te_var}, the variance of $\hat{\beta}_t^{\MTL}$, is always smaller than the variance of $\hat{\beta}_t^{\STL}$.
This part introduces a positive variance reduction effect to performing multi-task learning.
Hence, whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$ is determined by the trade-off between two effects:
(i) the positive effect from variance reduction;
(ii) the negative effect from model shift bias.

\todo{} We illustrate the example with a simulation.
We consider a setting where $p = 200$, $n_1 = 90p$, $n_2 = 30p$.
We fix the target task and vary the source task, in particular the parameter $d$ which determines $\norm{\beta_1 - \beta_2}$.
Figure \ref{fig_model_shift} shows the result.
We observe that Proposition \ref{prop_dist_transition} explains most of the observations in Figure \ref{fig_model_shift}.
%in this proposition (and the following ones) can be replaced with $(1+c_1^{-1/2})^{-4}$ and $(1-c_1^{-1/2})^{-4}$. In particular, these bounds becomes tighter as $c_1$ increases.

@@

\textbf{Cause 2.}
In classical Rademacher or VC based theory of multi-task learning, the generalization bounds are usually presented for settings where the data sizes are equal for all tasks \cite{B00,M06,MPR16}.
More generally, such results are still applicable when all task data are being added simultaneously.
On the other hand, for many applications of multi-task learning, the data sources are usually heterogeneous.
For such settings, we have observed that adding more labeled data from one task does not always help.
%On the other hand, we have observed that adding more labeled data does not always improve performance in multi-task learning.
Using the isotropic model, we consider what happens if we vary the source task data size.

@@

\textbf{Cause 3.}
So far we have considered the isotropic model where $\Sigma_1 = \Sigma_2$.
This setting is relevant for settings where different tasks share the same input features such as multi-class image classification.
In general, the covariance matrices of the two tasks may be different such as in text classification.
In this part, we consider what happens when $\Sigma_1 \neq \Sigma_2$.

@@

\subsection{Implications for Positive Transfer}\label{sec_benefit}

On the positive side, we describe two implications for positive transfer, assuming that $\Psi(\beta_1, \beta_2)$ is small in the isotropic model.

\textbf{Implication 1.}
We use our tools to explain a key result of taskonomy \cite{ZSSGM18}, which shows that by learning from multiple related tasks, one can reduce the amount of labeled data from each task.
This is formalized as follows.

\textbf{Implication 2.}
We further show that multi-task learning is particular powerful when the labeled data of the source task is less noisy compared to the target task.
Consider a more general setting of Proposition \ref{prop_dist_transition} where the noise level $\sigma_1$ of task $1$ differs from the noise level $\sigma_2$ of task $2$.
We extend Proposition \ref{prop_dist_transition} as follows.

>>>>

\section{Understanding Negative Transfer and Their Algorithmic Implications}\label{sec_insight}

Based on the key lemma, we explain three contributing factors of negative transfer, including model dissimilarity, large source data size and covariate shift.
For each factor, we show how negative transfer occurs using a simplified isotropic model.
Then, we describe the algorithmic implication of the theory.

\subsection{Model Dissimilarity}

@@

We measure model dissimilarity as $\norm{\beta_1 - \beta_2}^2$, which is the distance between source and target in the isotropic model.
We derive a sharp threshold when positive transfer transitions to negative transfer, as model dissimilarity increases.
%Based on Theorem \ref{thm_model_shift}, we derive the transition threshold in the following proposition.
Define the following functions
\begin{align*}
	\Psi(\beta_1, \beta_2) = {\ex{\bignorm{\beta_1 - \beta_2}^2}} / {\sigma^2},  \Phi(\rho_1, \rho_2) = \frac{(\rho_1 + \rho_2 - 1)^2}{\rho_1 (\rho_1 + \rho_2) (\rho_2 - 1)}.
\end{align*}
Let $\alpha_{-} = (1 - \rho_1^{-1/2})^4$ and $\alpha_{+} = (1 + \rho_1^{-1/2})^4$.

@@

Proposition \ref{prop_dist_transition} simplifies Theorem \ref{thm_main_informal} in the isotropic model, allowing for a more explicit statement of the bias-variance tradeoff.
\todo{} We illustrate the example with a simulation.
We consider a setting where $p = 200$, $n_1 = 90p$, $n_2 = 30p$.
We fix the target task and vary the source task, in particular the parameter $d$ which determines $\norm{\beta_1 - \beta_2}$.
Figure \ref{fig_model_shift} shows the result.
We observe that Proposition \ref{prop_dist_transition} explains most of the observations in Figure \ref{fig_model_shift}.

@@

\textbf{Algorithmic implication.}
We further show that multi-task learning is particular powerful when the labeled data of the source task is less noisy compared to the target task.
Consider a more general setting of Proposition \ref{prop_dist_transition} where the noise level $\sigma_1$ of task $1$ differs from the noise level $\sigma_2$ of task $2$.
If $\sigma_1$ is too large, then the source task provides a negative transfer to the target.
If $\sigma_1$ is small, then the transfer may be positive.
Proposition \ref{prop_var_transition} formalizes the intuition.

Based on the result, we propose a simple metric to determine whether multi-task learning performs better than single-task learning.
The metric is as follows.
First, we train 2 single-task models on 2 separate tasks.
We take the prediction accuracy of the two tasks.
Let $\tau$ be a fixed threshold.
If the accuracy of the source task is higher than the accuracy of the target task by $\tau$, then we predict positive transfer.
On the other hand, if the accuracy of the source task is lower than the accuracy of the target task by $\tau$, then we predict negative transfer.

\subsection{Data Size.}
In classical Rademacher or VC based theory of multi-task learning, the generalization bounds are usually presented for settings where the data sizes are equal for all tasks \cite{B00,M06,MPR16}.
More generally, such results are still applicable when all task data are being added simultaneously.
On the other hand, for many applications of multi-task learning, the data sources are usually heterogeneous.
For such settings, we have observed that adding more labeled data from one task does not always help.
%On the other hand, we have observed that adding more labeled data does not always improve performance in multi-task learning.
Using the isotropic model, we consider what happens if we vary the source task data size.

@@

\textbf{Algorithmic implication.}
We use our tools to explain a key result of taskonomy \cite{ZSSGM18}, which shows that by learning from multiple related tasks, one can reduce the amount of labeled data from each task.
This is formalized as follows.

\subsection{Covariate Shift}

So far we have considered the isotropic model where $\Sigma_1 = \Sigma_2$.
This setting is relevant for settings where different tasks share the same input features such as multi-class image classification.
In general, the covariance matrices of the two tasks may be different such as in text classification.
In this part, we consider what happens when $\Sigma_1 \neq \Sigma_2$.

@@

\textbf{Algorithmic implication.}
